---
title: "Explore qua-kit grading system"
author: "Artem Chirkin"
date: "July 27, 2017"
output: html_document
---

## Prepare datasets

First of all, I need to load libraries and set up PostgreSQL connection.
```{r loading libraries, echo=TRUE, results='hide', message=FALSE}
require(RPostgreSQL)
require(ggplot2)
require(gridExtra)
# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")
# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
con <- dbConnect(drv, dbname = "qua-kit-old-ratings")
```

The next step is to get all necessary data from DB.
```{r query data, echo=TRUE, results='hide', message=FALSE}
# all users
#db_users <- dbGetQuery(con, "SELECT id,name,role FROM \"user\"")

#db_scenario <- dbGetQuery(con, "SELECT id,author_id,task_id,last_update FROM scenario")

# all cross feedback
#db_criterion <- dbGetQuery(con, "SELECT * FROM criterion")
db_vote <- dbGetQuery(con, paste
    ( "SELECT w.task_id, v.id, v.criterion_id, v.voter_id, v.better_id, v.worse_id, b.last_update as better_time, w.last_update as worse_time, rb.value as better_value, rw.value as worse_value, b.author_id as better_author, w.author_id as worse_author, v.timestamp"
    , "  FROM vote v, rating rb, rating rw, scenario b, scenario w"
    , " WHERE v.better_id = b.id AND v.worse_id = w.id"
    , "   AND rw.author_id = w.author_id AND rb.author_id = b.author_id"
    , "   AND rw.problem_id = w.task_id AND rb.problem_id = b.task_id"
    , "   AND rw.criterion_id = v.criterion_id AND rb.criterion_id = v.criterion_id"
    , "   AND rw.value != 0 AND rb.value != 0"
#    , "   AND v.criterion_id = 4"
    ))
#db_review <- dbGetQuery(con, "SELECT * FROM review")

# cumulative
db_rating <- dbGetQuery(con, "SELECT rating.* FROM rating WHERE rating.value > 0")
db_vote_grade <- dbGetQuery(con, "SELECT * FROM vote_grade") # old table
#db_vote_rating <- dbGetQuery(con, "SELECT * FROM vote_rating")


# drop connection, since I don't need anything else in the DB.
dbDisconnect(con)
rm(con,drv)
```



```{r prepare data}
# make rating and vote table have normalized values (zero mean, 1 std)
rating <- db_rating
vote <- db_vote
vote_rating <- db_vote_grade
vote_rating$grade <- (db_vote_grade$grade - mean(db_vote_grade$grade))/sqrt(var(db_vote_grade$grade))
for (c_id in unique(db_rating$criterion_id) ) {
    rows <- which(db_rating$criterion_id == c_id)
    rowsv <- which(db_vote$criterion_id == c_id)
    m <- mean(db_rating[rows,"value"])
    v <- var(db_rating[rows,"value"])
    rating[rows,"value"] <- (db_rating[rows,"value"] - m)/sqrt(v)
    vote[rowsv,"better_value"] <- (db_vote[rowsv,"better_value"] - m)/sqrt(v)
    vote[rowsv,"worse_value"] <- (db_vote[rowsv,"worse_value"] - m)/sqrt(v)
}
# total number of votes
n <- nrow(vote)
rm(rows,rowsv,m,v,c_id)
```


## Plot old data


```{r}
qplot(value, data=rating, geom="histogram") + ggtitle("submission ratings histogram")
#qplot(grade, data=vote_rating, geom="histogram")

# qplot(value, data=db_rating[which(db_rating$value > 0 & db_rating$criterion_id == 1),], geom="histogram")
```

## modelling decisions

We can say that voter's decisions are totally random, because submissions i an j come from the same pool, they are indifferent.
We can draw decision value from Bernoulli distribution ($p = 0.5$), then reshuffle worse/better submissions in voting table.
Thus, the decisions become a proper random variable, even though highly correlated with submissions ratings.


```{r}
delta <- 2*rbinom(n,1,0.5)-1
rating_i <- mapply(function(d,vb,vw) if (d > 0) vb else vw, delta, vote$better_value, vote$worse_value )
rating_j <- mapply(function(d,vb,vw) if (d > 0) vw else vb, delta, vote$better_value, vote$worse_value )
rating_v <- sapply(vote$voter_id, function(i) vote_rating[which(vote_rating$student_id == i), "grade"][1])
rating_v[is.na(rating_v)] <- 0
# qplot( rating_j, geom="histogram")
z <- delta * exp( rating_v + delta * (rating_j - rating_i) )

Dz <- var(z)
Kz <- (cov(rating_i, z) - cov(rating_j, z))/2
```

## output values

```{r}
# covariance of rating i and choice
cov(rating_i, delta)

# covariance of rating j and choice
cov(rating_j, delta)

# variance of unknown random variable Z
Dz

# covariance of unknown random variable Z with r_i (or inverse of cov for r_j)
Kz
```


## Output plots

```{r}
worse <- qplot(worse_value, data=vote, geom="histogram") + ggtitle("Votes - worse")
better <- qplot(better_value, data=vote, geom="histogram") + ggtitle("Votes - better")
grid.arrange(worse, better, ncol=2)
rm(worse,better)

pi <- qplot(rating_i, geom="histogram") + ggtitle("Votes - i")
pj <- qplot(rating_j, geom="histogram") + ggtitle("Votes - j")
grid.arrange(pi, pj, ncol=2)
rm(pi, pj)
```
